---
title: "Jonasobble"
output: html_notebook
---

# Load Libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(primes)
library(httr)
library(jsonlite)
library(magick)

library(packcircles) # to distribute circles (representing images on the canvas)
library(ggforce) # for drawing outer circle
library(ggimage) # for including images in ggplot
```



# Introduction

Dobble is based on the freaky fun of [finite](https://en.wikipedia.org/wiki/Finite_geometry) [projective](https://en.wikipedia.org/wiki/Projective_geometry) geometry.  
Exact explanation is a hard TBD.  

For now, here's some good resources to understand it (ordered by level of brainfuck):

1. https://puzzlewocky.com/games/the-math-of-spot-it/
2. http://www.petercollingridge.co.uk/blog/mathematics-toys-and-games/dobble/
3. http://www.mathpuzzle.com/MAA/47-Fano/mathgames_05_30_06.html

Following is a list of orders for which we can construct Dobble sets, the number of symbols necessary to do so and the number of cards generated (equal to the number of symbols on each card).

```{r}
projective_plane_orders <- tibble(
    order = sort(c(
      primes::generate_primes(min=1, max=100),
      primes::generate_primes(min=1, max=50)^2,
      primes::generate_primes(min=1, max=20)^3,
      primes::generate_primes(min=1, max=20)^4
    )),
    number_of_cards = order^2 + order + 1,
    number_of_symbols = number_of_cards,
    symbols_per_card = order + 1
  )

projective_plane_orders
```



# Init Project

## Set Params

```{r}
# backfill settings
pexels_api_key <- jsonlite::read_json('keys.json')[['pexels_api_key']]

# order of the projective geometry, governing the number of cards generated & images needed
game.order <- 7

pdf.nrow <- 3
pdf.ncol <- 2


image_resize.max_side <- 800

backfill_images <- TRUE
```

```{r}
game.number_of_cards <- projective_plane_orders %>% filter(order == game.order) %>% pull(number_of_cards)
game.number_of_symbols <- projective_plane_orders %>% filter(order == game.order) %>% pull(number_of_symbols)
game.symbols_per_card <- projective_plane_orders %>% filter(order == game.order) %>% pull(symbols_per_card)

pdf.cards_per_page <- pdf.nrow * pdf.ncol
```

All set for a game of *J(onas)obble* of order `r game.order`!  
This will generate a set of `r game.number_of_cards` cards, with `r game.symbols_per_card` symbols each.  
(For this we'll need `r game.number_of_symbols` in total, provided into the `/images/input/` folder. If there are less, we'll try to backfill with cat pictures from [pexels.com](https://pexels.com).


## Prepare images

### Set folder structure

```{r}
image_folder_path <- 'images'
output_folder_path <- 'card_output'

if(!dir.exists(image_folder_path)) { dir.create(image_folder_path) }

image_folder_path.input <- file.path(image_folder_path, 'input')
if(!dir.exists(image_folder_path.input)) { dir.create(image_folder_path.input) }

image_folder_path.backfill <- file.path(image_folder_path, 'backfill')
if(!dir.exists(image_folder_path.backfill)) { dir.create(image_folder_path.backfill) }

image_folder_path.final <- file.path(image_folder_path, 'final')
if(!dir.exists(image_folder_path.final)) { dir.create(image_folder_path.final) }

if(!dir.exists(output_folder_path)) { dir.create(output_folder_path) }
```


### Backfill images

If not enough images are provided yet, backfill from Pexels.com (make sure to drop a valid API key into `keys.json`)

```{r warning=FALSE}
# fill the dummy folder
image_files.input <- list.files(path = image_folder_path.input, pattern = '\\.(jpg|jpeg|png)$', full.names = T, recursive = F)
image_files.backfill <- list.files(path = image_folder_path.backfill, pattern = '\\.(jpg|jpeg|png)$', full.names = T, recursive = F)

if(backfill_images & !is_empty(pexels_api_key)) {
  backfill.batch_size <- 15

  backfill.max_batches <- ceiling(game.number_of_symbols/backfill.batch_size)
  backfill.num_images_to_fill <- max(game.number_of_symbols - length(image_files.input) - length(image_files.backfill), 0)
  backfill.num_pages_to_query <- ceiling(backfill.num_images_to_fill / backfill.batch_size)
  
  if(backfill.num_pages_to_query > 0) {
    for(batch_i in tail(1:backfill.max_batches, backfill.num_pages_to_query)) {
      backfill_images_req <- httr::GET(
        paste0('https://api.pexels.com/v1/search?query=cat&per_page=',backfill.batch_size,'&page=',batch_i),
        add_headers(Authorization = pexels_api_key)
      )
      
      backfill_images <- content(backfill_images_req)$photos %>%
        map_dfr(as.data.frame)
      
      for(i in 1:nrow(backfill_images)) {
        download.file(
          url = ifelse(
            backfill_images$height[i] > 2000 | backfill_images$width[i] > 2000,
            backfill_images$src.large[i],
            backfill_images$src.original[i]
          ),
          destfile = file.path(image_folder_path.backfill, basename(backfill_images$src.original[i])),
          mode = 'wb',
          #quiet = T
        )
        Sys.sleep(0.2)
      }
      
    }
  }
  
  image_files.backfill <- list.files(path = image_folder_path.backfill, pattern = '\\.(jpg|jpeg|png)$', full.names = T, recursive = F)
}


image_files.src <- tibble(
    image_file_path = c(image_files.input, image_files.backfill),
    image_type = c(rep('input', length(image_files.input)), rep('backfill', length(image_files.backfill)))
  ) %>%
  bind_cols(
    magick::image_read(.$image_file_path) %>% magick::image_info()
  )

image_files.final <- image_files.src %>%
  select (
    image_file_path.src = image_file_path, image_type
  )

for(i in 1:nrow(image_files.final)) {
  tryCatch({
    img <- magick::image_read(image_files.final$image_file_path.src[i])
    img.resized <- image_resize(img, paste(image_resize.max_side,image_resize.max_side,sep='x'))
    
    image_file_path.resized <- file.path(image_folder_path.final, basename(image_files.final$image_file_path.src[i]))
    
    magick::image_write(img.resized, image_file_path.resized)
    print(sprintf('resized %s to the dimensions %sx%s px', basename(image_files.final$image_file_path.src[i]), magick::image_info(img.resized)$width, magick::image_info(img.resized)$height))
    
    image_files.final$image_file_path.resized[i] <- image_file_path.resized
  })
}

image_files.final <- image_files.final %>%
  filter(!is.na(image_file_path.resized)) %>%
  bind_cols(
    magick::image_read(.$image_file_path.resized) %>% magick::image_info()
  )
```

We've now gathered & resized `r nrow(image_files.final)` image files to work into the game


# Generator functions

## Dobble Permutations

This function will return a list of symbol permutations (representing individual game-cards) for a given `order`.

```{r}
# taken code from https://math.stackexchange.com/a/1956107
generate_dobble_permutations <- function(order) {
  card_symbols <- list()
  
  # build the initial set of cards as per 
  for(i in 0:order) {
    symbol_indices <- c(1, 1:order + i*order + 1)
    
    card_symbols[[paste(1,i+1,sep='_')]] <- symbol_indices
  }
  
  # build the rest of the deck
  for(i in seq(0,order-1)) {
    for(j in seq(0,order-1)) {
      k <- seq(0,order-1)
      symbol_indices <- c(i+1+1, order+1 + order*k +(i*k+j)%%order +1)
      card_symbols[[paste(i+2,j+1,sep='_')]] <- symbol_indices
    }
  }
  
  return(card_symbols)
}
```


## Distribute Images

This function generates coordinates on & within a `radius = 1` circle where we can position our images.  
Based on the patterns of [sunflower seeds](https://momath.org/home/fibonacci-numbers-of-sunflower-seed-spirals/).

```{r}
# https://stackoverflow.com/a/28572551/1335174
generate_image_coordinates.sunflower <- function(n, alpha = 2, geometry = c('planar','geodesic')) {
  radius <- function(k,n,b) {
    ifelse(k > n-b, 1, sqrt(k-1/2)/sqrt(n-(b+1)/2))
  }
  
  b <- round(alpha*sqrt(n))  # number of boundary points
  phi <- (sqrt(5)+1)/2  # golden ratio
  
  r <- radius(1:n,n,b)
  #theta <- 2*pi*1:n/phi^2 
  theta <- 1:n * ifelse(geometry[1] == 'geodesic', 360*phi, 2*pi/phi^2)
  
  tibble(
    symbol_on_card = seq(n),
    x = r*cos(theta),
    y = r*sin(theta)
  )
}

generate_image_coordinates.circlepacking <- function(image_file_paths, margin = 0) {
  image_data <- magick::image_read(image_file_paths) %>%
    magick::image_info() %>%
    mutate(
      image_file_path = image_file_paths,
      diagonal = sqrt(height^2 + width^2),
      radius = diagonal/2 + margin, # add a margin to the radius
    ) %>%
    arrange(desc(radius))
  
  image_data %>%
    select(image_file_path, height, width, diagonal, radius) %>%
    bind_cols(
      packcircles::circleProgressiveLayout(.$radius, sizetype = 'radius')
    ) %>%
    mutate(
      radius = diagonal/2
    )
}

```


# Create Dobble cards

## Calculate Symbol Permutations

```{r}
dobble_cards.permutations <- generate_dobble_permutations(game.order)
#dobble_cards.images <- lapply(dobble_cards.permutations, function(indices){image_files.final[indices]})

#dobble_images.matrix <- matrix(nrow = length(dobble_images), ncol = length(dobble_images), dimnames = list(dobble_images,dobble_images))
#for(i in 1:length(dobble_cards.permutations)) {
#  dobble_images.matrix[i,dobble_cards.permutations[[i]]] <- 1
#}

dobble_cards.df <- dobble_cards.permutations %>%
  map_dfr(.f = function(indices){
    tibble(
      symbol = indices,
      symbol_on_card = seq(symbol)
    )},
    .id = 'card'
  ) %>%
  mutate(
    card_generation_block = str_extract(card, '^\\d+(?=_)'),
    card_id = as.integer(as_factor(card)),
    image_file_path = image_files.final$image_file_path.resized[symbol],
    image_name = basename(image_file_path)
  )
```

We can plot the distribution of images (x-axis) on our cards (y-axis):

```{r fig.height=7, fig.width=7}
dobble_cards.df %>%
  ggplot() +
  geom_raster(
    aes(
      x = symbol,
      y = fct_rev(as_factor(card)),
      fill = card_generation_block
    )
  ) +
  scale_color_discrete(aesthetics = c('fill','color'), guide = F) +
  coord_fixed() +
  labs(
    x = 'symbol',
    y = 'card'
  )
```


## Distribute Images on Cards

```{r}
image_margin <- image_resize.max_side * 0.025 # for 1000px resized images, this is 5px each, so 10px spacing in between

# calculate the image coordinates for each card
dobble_cards.image_coords <- dobble_cards.df %>%
  group_by(card_id) %>%
  group_map(~generate_image_coordinates.circlepacking(.$image_file_path, margin = image_margin)) %>% # adds a margin of 5px to space the images a little
  # unfortunately, the packcircles methods don't center the results around (0,0), so we correct their x & y values here
  mutate(
    tmp.real_circle_center.x = (min(x-radius) + max(x+radius))/2,
    tmp.real_circle_center.y = (min(y-radius) + max(y+radius))/2,
    
    x = x - tmp.real_circle_center.x,
    y = y - tmp.real_circle_center.y
  ) %>%
  ungroup() %>%
  select(-starts_with('tmp.'))

dobble_cards.card_radius <- dobble_cards.image_coords %>%
  summarise(
    outer_radius = max(sqrt(x^2+y^2) + radius)
  ) %>%
  pull(outer_radius) %>%
  ceiling() %>%
  {. + 2}
```

We've laid out all of the images on our cards, given them a margin of `r image_margin` px each, so that they would have a little space in between them.  
To accommodate all pictures safely, the cards will all have a radius of `r dobble_cards.card_radius` px.

```{r}
dobble_cards.image_boundaries.circles <- dobble_cards.image_coords %>%
  mutate(
    id = paste(card_id, image_file_path, sep = '_')
  ) %>%
  select(x, y, radius, id) %>%
  packcircles::circleLayoutVertices(idcol = 'id') %>%
  mutate(
    card_id = as.integer(str_extract(id, '^\\d+(?=_)')),
    image_file_path = str_extract(id, '(?<=^\\d{1,5}_).+$')
  )

dobble_cards.image_boundaries.squares <- dobble_cards.image_coords %>%
  mutate(
    x0 = x - 0.5*width,
    x1 = x + 0.5*width,
    y0 = y - 0.5*height,
    y1 = y + 0.5*height
  )
```

We can plot this without the actual images, to see a bit more clearly how this will look like for the first 4 cards.

```{r fig.height=9, fig.width=7}
ggplot() +
  geom_polygon(
    data = dobble_cards.image_boundaries.circles %>%
      filter(card_id <= pdf.cards_per_page),
    aes(
      x, y,
      group = id,
      fill = image_file_path
    ),
    color = 'grey25',
    alpha = 0.5
  ) +
  geom_rect(
    data = dobble_cards.image_boundaries.squares %>%
      filter(card_id <= pdf.cards_per_page),
    aes(
      xmin = x0, xmax = x1,
      ymin = y0, ymax = y1,
      fill = image_file_path
    ),
    color = 'grey50',
    alpha = 0.8
  ) +
  geom_text(
    data = dobble_cards.image_coords %>%
      filter(card_id <= pdf.cards_per_page) %>%
      left_join(dobble_cards.df, by = c('card_id', 'image_file_path')),
    aes(
      x, y,
      label = symbol
    )
  ) +
  ggforce::geom_circle(
    data = tibble(card_id = 1:pdf.cards_per_page),
    aes(
      x0 = 0,
      y0 = 0,
      r = dobble_cards.card_radius
    ),
    fill = 'blue',
    alpha = 0.1
  ) +
  scale_color_discrete(aesthetics = c('color','fill'), guide = F) +
  coord_fixed() +
  facet_wrap(~card_id, nrow = pdf.nrow, ncol = pdf.ncol) +
  theme(panel.spacing=unit(0, 'mm'))
```


## Finalize Card Set

We'll bind the calculated image coordinates to our previously generated permutation data and receive a data table that tells us exactly which image will go on which card and where.

```{r}
dobble_cards.final <- dobble_cards.df %>%
  left_join(
    dobble_cards.image_coords,
    by = c('card_id', 'image_file_path')
  )
```

Using that, we'll generate a pdf with `r ceiling(game.number_of_cards/pdf.cards_per_page)` pages (holding `r pdf.cards_per_page` cards each) into the output folder `r output_folder_path`.

```{r}
pdf_pages <- tibble(
    page_i = 1:ceiling(game.number_of_cards/pdf.cards_per_page),
    pdf_content.cards.first_id = (page_i-1)*pdf.cards_per_page +1,
    pdf_content.cards.last_id = pmin(page_i*pdf.cards_per_page, game.number_of_cards)
  )


for(page_i in 1:nrow(pdf_pages)) {
  pdf_page <- pdf_pages[page_i,]
  
  pdf_file_path = file.path(output_folder_path, paste(strftime(lubridate::now(), '%Y-%m-%d'), sprintf('jonasobble_page_%02d.pdf',page_i), sep = '_'))
  pdf_title = sprintf('J(onas)obble Game, page %s of %s', page_i, nrow(pdf_pages))
  
  #pdf(
  #  file = pdf_file_path,
  #  title = pdf_title,
  #  paper = 'a4',
  #  width = 8, height = 11.5,
  #  bg = 'transparent',
  #  compress = F
  #)
  
  pdf_page_plot <- ggplot() +
    ggforce::geom_circle(
      data = tibble(card_id = pdf_page$pdf_content.cards.first_id:pdf_page$pdf_content.cards.last_id),
      aes(
        x0 = 0,
        y0 = 0,
        r = dobble_cards.card_radius
      ),
      #fill = 'blue',
      #alpha = 0.2
    ) +
    ggimage::geom_image(
      data = dobble_cards.final %>%
        filter(
          card_id >= pdf_page$pdf_content.cards.first_id,
          card_id <= pdf_page$pdf_content.cards.last_id
        ),
      aes(
        x = x,
        y = y,
        image = image_file_path,
        size=I(0.5*width/dobble_cards.card_radius)
      )
    ) +
    coord_fixed() +
    facet_wrap(~card_id, nrow = pdf.nrow, ncol = pdf.ncol, strip.position = 'left') +
    theme_void()
  
  #print(pdf_page_plot)
  ggsave(
    filename = paste(strftime(lubridate::now(), '%Y-%m-%d'), sprintf('jonasobble_page_%02d.png',page_i), sep = '_'),
    device = 'png',
    width = 210, height = 297, units = 'mm', # a4
    dpi = 320,
    path = output_folder_path
  )
  
  #dev.off()
  print(paste('printed page',page_i))
}
```

